IFC Inside: A General Approach to Retrofitting Languages with Dynamic Information Flow Control

### 2015-01-07 paper/appendix.tex
Figure -> Fig.

In Figure~\ref{fig:ml-full} we give the full semantics for |targetLangML|.  A subset
of them has been given in Figure~\ref{fig:ml} earlier in the paper.

@@

The essential parts of the semantics for the concrete language
with a single heap,
which we call |concreteLangMl alpha|,
are given in Figure~\ref{fig:concrete}.  Most rules are
straight-forward translations of the rules in Figures~\ref{fig:ifc}
and~\ref{fig:embedding} but for a single heap.  For conciseness, we
only show the interesting ones.
Now, we can show an isomorphism between this language and
|restrictedLangNoRef alpha targetLangML|, which
(by Theorem~\ref{thm:iso-tsni} and~\ref{thm:iso-tini}) guarantees
non-interference for an appropriate scheduling policy |alpha|.

@@

In Figure~\ref{fig:labeled-vals}, we formally show how to add this
extension to the language.
We assume that the constructor |Labeled| is not part
of the surface syntax, but rather an internal construct.

@@

In Figure~\ref{fig:labeled-refs} details labeled references formally.
The construct |ial| is internal in the labeled reference implementation,
and not part of the surface syntax.
The changes to the language for labeled values and
references require us to update the erasure function
|erasef il|, whose full definition is shown in Figure~\ref{fig:erasure2}.

@@

\begin{figure}
  \begin{align*}
  &|erase il (iconf iS its) =
  iconf (erase il iS) (fltr (\ it . it = bullet) (map (erasef il) its))|\\
  &|fullconf id il' tS ie| \begin{cases}
  |bullet| & |il' cantFlowTo il| \\
  |fullconf id il' (erase il tS) (erase il ie)| & \text{otherwise}
  \end{cases} \\
  &|erase il (Labeled il' ie)|= \begin{cases}
  |Labeled il' bullet| & |il' cantFlowTo il| \\
  |Labeled il' ie| & \text{otherwise}
  \end{cases} \\
  &|erase il emptyset = emptyset|\\
  &|erase il (bracket iS (mapsto id Q)) =| \begin{cases}
  |erase il iS| & \text{|il' cantFlowTo il|, where |il'| is the label of thread |id|}\\
  |bracket (erase il iS) (mapsto id (erase il Q))| & \text{otherwise}
  \end{cases} \\
  &|erase il (bracket iS (mapsto (iall il') iv)) =| \begin{cases}
  |bracket (erase il iS) (mapsto (iall il') bullet)| & \text{|il' cantFlowTo il|}\\
  |bracket (erase il iS) (mapsto (iall il') (erase il iv))| & \text{otherwise}
  \end{cases} \\
  &|erase il Q = filter il Q|
  \end{align*}
  \caption{Erasure function for the full IFC language, with all extensions.
    In all cases that are not specified, including target-language constructs,
    |erasef il| is applied homomorphically
    (e.g., |erase il (setLabel e) = setLabel (erase il e)|).
    This definition replaces the one from Figure~\ref{fig:erasure}, which
    is for the IFC language without extensions.}
  \label{fig:erasure2}
\end{figure}

>>>>

In Fig.~\ref{fig:ml-full} we give the full semantics for |targetLangML|.  A subset
of them has been given in Fig.~\ref{fig:ml} earlier in the paper.

@@

The essential parts of the semantics for the concrete language
with a single heap,
which we call |concreteLangMl alpha|,
are given in Fig.~\ref{fig:concrete}.  Most rules are
straight-forward translations of the rules in Figs.~\ref{fig:ifc}
and~\ref{fig:embedding} but for a single heap.  For conciseness, we
only show the interesting ones.
Now, we can show an isomorphism between this language and
|restrictedLangNoRef alpha targetLangML|, which
(by Theorem~\ref{thm:iso-tsni} and~\ref{thm:iso-tini}) guarantees
non-interference for an appropriate scheduling policy |alpha|.

@@

In Fig.~\ref{fig:labeled-vals}, we formally show how to add this
extension to the language.
We assume that the constructor |Labeled| is not part
of the surface syntax, but rather an internal construct.

@@

In Fig.~\ref{fig:labeled-refs} details labeled references formally.
The construct |ial| is internal in the labeled reference implementation,
and not part of the surface syntax.
The changes to the language for labeled values and
references require us to update the erasure function
|erasef il|, whose full definition is shown in Fig.~\ref{fig:erasure2}.

@@

\begin{figure}
  \begin{align*}
  &|erase il (iconf iS its) =
  iconf (erase il iS) (fltr (\ it . it = bullet) (map (erasef il) its))|\\
  &|fullconf id il' tS ie| \begin{cases}
  |bullet| & |il' cantFlowTo il| \\
  |fullconf id il' (erase il tS) (erase il ie)| & \text{otherwise}
  \end{cases} \\
  &|erase il (Labeled il' ie)|= \begin{cases}
  |Labeled il' bullet| & |il' cantFlowTo il| \\
  |Labeled il' ie| & \text{otherwise}
  \end{cases} \\
  &|erase il emptyset = emptyset|\\
  &|erase il (bracket iS (mapsto id Q)) =| \begin{cases}
  |erase il iS| & \text{|il' cantFlowTo il|, where |il'| is the label of thread |id|}\\
  |bracket (erase il iS) (mapsto id (erase il Q))| & \text{otherwise}
  \end{cases} \\
  &|erase il (bracket iS (mapsto (iall il') iv)) =| \begin{cases}
  |bracket (erase il iS) (mapsto (iall il') bullet)| & \text{|il' cantFlowTo il|}\\
  |bracket (erase il iS) (mapsto (iall il') (erase il iv))| & \text{otherwise}
  \end{cases} \\
  &|erase il Q = filter il Q|
  \end{align*}
  \caption{Erasure function for the full IFC language, with all extensions.
    In all cases that are not specified, including target-language constructs,
    |erasef il| is applied homomorphically
    (e.g., |erase il (setLabel e) = setLabel (erase il e)|).
    This definition replaces the one from Fig.~\ref{fig:erasure}, which
    is for the IFC language without extensions.}
  \label{fig:erasure2}
\end{figure}

### 2015-01-07 paper/formal.tex
Figure -> Fig.

%{
%format .->  = "\overset{\alpha}{\hookrightarrow}"
Our language |specLang alpha (targetLang nop)| fulfills
this definition as |(ifc C, .->, erasef il)|, where
$|ifc C| = |iS| \times \operatorname{List}(|it|)$.  The set of terminal conditions
$|ifc V|$ is $|iS| \times |it|_V$, where $|it|_V \subset |it|$ is the
type for tasks whose expressions have been reduced to
values.\footnote{
  Here, we abuse notation by describing types for configuration parts using the
  same metavariables as the ``instance'' of the type, e.g., |it| for the type of
  task.
}
The erased configuration |tyerase (ifc C)| extends |ifc C| with configurations
containing |bullet|, and Figure~\ref{fig:erasure} gives the precise definition for
our erasure function |erasef il|.
%
Essentially, a task and its corresponding message queue is completely erased from the task
list if its label does not flow to the attacker observation level |il|.
Otherwise, we apply the erasure function homomorphically and remove any message
from the task's message queue that are more sensitive than |il|.
%}

@@

Our IFC language satisfies TSNI %under a particular
%scheduling policy.
%
%Specifically, we consider the language
under the round-robin scheduler
|roundrobinf| given in Figure~\ref{fig:scheduler}.

@@

However, even non-adversarial schedulers are not always safe.
Consider, for example, the sequential scheduling policy |seqf| given in
Figure~\ref{fig:scheduler}.
%
It is easy to show that |specLang seqf targetLang| does not satisfy
TSNI:
%
consider a target language similar to |targetLangML| with an
additional expression terminal |diverge| that denotes a divergent computation,
i.e., |diverge| always reduces to |diverge| and a simple label lattice |{pub,
sec}| such that |canFlowTo2 pub sec|, but |cantFlowTo2 sec pub|.
Consider the following two configurations in this language:
\begin{code}
ic1 = iS ; fullconf 1 sec tS1 (IT(if false then diverge else true)),  fullconf 2 pub tS2 ie
ic2 = iS ; fullconf 1 sec tS1 (IT(if true then diverge else true)),   fullconf 2 pub tS2 ie
\end{code}
These two configurations are |pub|-equivalent, but |ic1| will reduce
(in two steps)
to |ic1 = iS ; fullconf 2 pub tS1 (IT true)|, whereas |ic2| will not make
any progress.
%
Suppose that |ie| is a computation that writes to a |pub| channel,\footnote{
Though we do not model labeled channels, extending the calculus with such a
feature is straight forward, see Section~\ref{sec:extensions}.}
then the |sec| task's decision to diverge or not is directly leaked to a
public entity.

>>>>

%{
%format .->  = "\overset{\alpha}{\hookrightarrow}"
Our language |specLang alpha (targetLang nop)| fulfills
this definition as |(ifc C, .->, erasef il)|, where
$|ifc C| = |iS| \times \operatorname{List}(|it|)$.  The set of terminal conditions
$|ifc V|$ is $|iS| \times |it|_V$, where $|it|_V \subset |it|$ is the
type for tasks whose expressions have been reduced to
values.\footnote{
  Here, we abuse notation by describing types for configuration parts using the
  same metavariables as the ``instance'' of the type, e.g., |it| for the type of
  task.
}
The erased configuration |tyerase (ifc C)| extends |ifc C| with configurations
containing |bullet|, and Fig.~\ref{fig:erasure} gives the precise definition for
our erasure function |erasef il|.
%
Essentially, a task and its corresponding message queue is completely erased from the task
list if its label does not flow to the attacker observation level |il|.
Otherwise, we apply the erasure function homomorphically and remove any message
from the task's message queue that are more sensitive than |il|.
%}

@@

Our IFC language satisfies TSNI %under a particular
%scheduling policy.
%
%Specifically, we consider the language
under the round-robin scheduler
|roundrobinf| given in Fig.~\ref{fig:scheduler}.

@@

However, even non-adversarial schedulers are not always safe.
Consider, for example, the sequential scheduling policy |seqf| given in
Fig.~\ref{fig:scheduler}.
%
It is easy to show that |specLang seqf targetLang| does not satisfy
TSNI:
%
consider a target language similar to |targetLangML| with an
additional expression terminal |diverge| that denotes a divergent computation,
i.e., |diverge| always reduces to |diverge| and a simple label lattice |{pub,
sec}| such that |canFlowTo2 pub sec|, but |cantFlowTo2 sec pub|.
Consider the following two configurations in this language:
\begin{code}
ic1 = iS ; fullconf 1 sec tS1 (IT(if false then diverge else true)),  fullconf 2 pub tS2 ie
ic2 = iS ; fullconf 1 sec tS1 (IT(if true then diverge else true)),   fullconf 2 pub tS2 ie
\end{code}
These two configurations are |pub|-equivalent, but |ic1| will reduce
(in two steps)
to |ic1 = iS ; fullconf 2 pub tS1 (IT true)|, whereas |ic2| will not make
any progress.
%
Suppose that |ie| is a computation that writes to a |pub| channel,\footnote{
Though we do not model labeled channels, extending the calculus with such a
feature is straight forward, see Section~\ref{sec:extensions}.}
then the |sec| task's decision to diverge or not is directly leaked to a
public entity.

### 2015-01-07 paper/real.tex
Figure -> Fig.

\begin{figure}[t]
\centerline{\includegraphics[width=0.7\columnwidth]{figs/node}}
\caption{\label{fig:node}
This example shows how our trusted monitor (left) is used to mediate
communication between two tasks for which IFC is enforced (right).}
\end{figure}
%
\paragraph{Server-side IFC for Node.js:}
We have implemented |specLangJS seqf| for Node.js in the form of a
library, without modifying Node.js or the V8 JavaScript engine.
%
Our implementation\footnote{Available at \codelink{}.} provides a
library for creating new tasks, i.e., contexts whose global object
only contains the standard JavaScript library and our IFC primitives
(e.g., |send| and |sandbox|).
%
When mapped to our formal treatement, |sandbox| is defined with |ap
klone tS = tS0|, where |tS0| is the global object corresponding to the
standard JavaScript library and our IFC primitives.
%
%These tasks can communicate and create new tasks, in turn.
%
These IFC operations are mediated by the trusted library code (executing
as the main Node.js context), which tracks the state (current label, messages,
etc.) of each task.  An example for |send|/|recv| is shown in
Figure~\ref{fig:node}.
%While V8 provides a means for sharing any object between contexts (used
%by same-origin contexts in Blink),
We remark that our system conservatively restricts
the kinds of messages that can be exchanged, via |send| (and |sandbox|),
to string values.
%
(We provide a convenience library which marshalls JSON objects to/from strings.)
%
In our formalization, this amounts to restricting the IFC language rule
for |send| in the following way:
%{
\newcommand{\str}{"string"}
%format tOf (e) = "\texttt{typeOf}("e")\texttt{ === \str}"
%format ttrue = "\texttt{true}"
\begin{mathpar}
\inferrule[JS-send]
{
|il canFlowTo il'|\\
|iS(id') = Q|\\
|iS' = iS [ mapsto id' (il', id,  iv) , Q ]|\\
|ie = IT te|\\
|conf tS (tOf te) -> conf tS ttrue|
}
{|
iconf iS (fullconf id il tS (iniEi (send id' il' iv)), ldots)
.->
iS'; sched step (fullconf id il tS (iniEi unit), ldots)
|}
\end{mathpar}
%}
%

While the described system implements |specLangJS seqf|, applications
typically require access to libraries (e.g., the file system library
\textsf{fs}) that have external effects.
%
Exposing the Node.js APIs directly to sandboxed tasks is unsafe.
Instead, we implement libraries (like a labeled version of \textsf{fs}) as
message exchanges between the sandboxed tasks (e.g., \textsf{task-1}
in Figure~\ref{fig:node}) and the main Node.js task that implements
the IFC monitor.
%
While this is safer than simply wrapping unsafe objects, which can
potentially be exploited to access objects outside the context, adding
features such as the \textsf{fs} requires the code in the main task to
ensures that labels are properly propagated and enforced.
%
Unfortunately, while imposing such a proof burden
is undesirable, this also has to be expected:
different language environments expose different libraries for
handling external I/O, and the correct treatment of external effects
is application specific.
%
%% For instance, in the case of Node.js, how labels interact with the
%% file system may even vary according to the application (e.g.,
%% in the IFC system Hails~\cite{hails}, labels are not persisted).
%
%While we can extend our formalism to model a particular interface to
%the file system, HTTP client, etc. the utility of this formalization is
%unclear and left to future work.
%
We do not extend our formalism to account for the  particular
interface to the file system, HTTP client, etc., as this is
specific to the Node.js implementation and does not generalize
to other systems.

>>>>

\begin{figure}[t]
\centerline{\includegraphics[width=0.7\columnwidth]{figs/node}}
\caption{\label{fig:node}
This example shows how our trusted monitor (left) is used to mediate
communication between two tasks for which IFC is enforced (right).}
\end{figure}
%
\paragraph{Server-side IFC for Node.js:}
We have implemented |specLangJS seqf| for Node.js in the form of a
library, without modifying Node.js or the V8 JavaScript engine.
%
Our implementation\footnote{Available at \codelink{}.} provides a
library for creating new tasks, i.e., contexts whose global object
only contains the standard JavaScript library and our IFC primitives
(e.g., |send| and |sandbox|).
%
When mapped to our formal treatement, |sandbox| is defined with |ap
klone tS = tS0|, where |tS0| is the global object corresponding to the
standard JavaScript library and our IFC primitives.
%
%These tasks can communicate and create new tasks, in turn.
%
These IFC operations are mediated by the trusted library code (executing
as the main Node.js context), which tracks the state (current label, messages,
etc.) of each task.  An example for |send|/|recv| is shown in
Fig.~\ref{fig:node}.
%While V8 provides a means for sharing any object between contexts (used
%by same-origin contexts in Blink),
We remark that our system conservatively restricts
the kinds of messages that can be exchanged, via |send| (and |sandbox|),
to string values.
%
(We provide a convenience library which marshalls JSON objects to/from strings.)
%
In our formalization, this amounts to restricting the IFC language rule
for |send| in the following way:
%{
\newcommand{\str}{"string"}
%format tOf (e) = "\texttt{typeOf}("e")\texttt{ === \str}"
%format ttrue = "\texttt{true}"
\begin{mathpar}
\inferrule[JS-send]
{
|il canFlowTo il'|\\
|iS(id') = Q|\\
|iS' = iS [ mapsto id' (il', id,  iv) , Q ]|\\
|ie = IT te|\\
|conf tS (tOf te) -> conf tS ttrue|
}
{|
iconf iS (fullconf id il tS (iniEi (send id' il' iv)), ldots)
.->
iS'; sched step (fullconf id il tS (iniEi unit), ldots)
|}
\end{mathpar}
%}
%

While the described system implements |specLangJS seqf|, applications
typically require access to libraries (e.g., the file system library
\textsf{fs}) that have external effects.
%
Exposing the Node.js APIs directly to sandboxed tasks is unsafe.
Instead, we implement libraries (like a labeled version of \textsf{fs}) as
message exchanges between the sandboxed tasks (e.g., \textsf{task-1}
in Fig.~\ref{fig:node}) and the main Node.js task that implements
the IFC monitor.
%
While this is safer than simply wrapping unsafe objects, which can
potentially be exploited to access objects outside the context, adding
features such as the \textsf{fs} requires the code in the main task to
ensures that labels are properly propagated and enforced.
%
Unfortunately, while imposing such a proof burden
is undesirable, this also has to be expected:
different language environments expose different libraries for
handling external I/O, and the correct treatment of external effects
is application specific.
%
%% For instance, in the case of Node.js, how labels interact with the
%% file system may even vary according to the application (e.g.,
%% in the IFC system Hails~\cite{hails}, labels are not persisted).
%
%While we can extend our formalism to model a particular interface to
%the file system, HTTP client, etc. the utility of this formalization is
%unclear and left to future work.
%
We do not extend our formalism to account for the  particular
interface to the file system, HTTP client, etc., as this is
specific to the Node.js implementation and does not generalize
to other systems.

### 2015-01-07 paper/retrofit.tex
Figure -> Fig.

In Figure~\ref{fig:ml}, we give a simple, untyped lambda calculus with
mutable references and fixpoint, prepared for combination with an
information flow control language.  The presentation is mostly standard, and utilizes Felleisen-Hieb reduction
semantics~\cite{Felleisen:1992:RRS:136293.136297} to define the
operational semantics of the system.  One peculiarity is that our language
defines an evaluation context |tE|, but, the evaluation rules have been
expressed in terms of a different evaluation context |curlyRedE|;
Here, we follow the approach of Matthews and
Findler~\cite{Matthews:2007:OSM:1190216.1190220} in order to simplify combining
semantics of multiple languages.
To derive the usual operational semantics for this language, the evaluation
context merely needs to be defined as |conf tS te @= tS, bracket tE te|.
However, when we combine this language with an IFC language, we
reinterpret the meaning of this evaluation context.

@@

In Figure~\ref{fig:ifc}, we give the syntax and \emph{single-task}
evaluation rules for a minimal information-flow control language.
Ordinarily, information flow control languages are defined by directly
stating a base language plus information flow control operators.  In
contrast, our language is purposely minimal: it does not have sequencing
operations, control flow, or other constructs.  However, it contains
support for the following core information flow control features:

@@

\subsection{The Embedding}
Figure~\ref{fig:embedding}
provides all of the rules responsible for actually carrying out the embedding of the IFC language with the target language.
The most important feature of this embedding is that every task maintains its own
copy of the target language global state and evaluation context, thus
enforcing isolation between various tasks.  In more detail:

@@

\begin{itemize}
    \item We extend the values, expressions and evaluation contexts of
      both languages to allow for embeddings terms in one language to
      be embedded in the other, as
      in~\cite{Matthews:2007:OSM:1190216.1190220}.
    %
In the target language, an IFC expression appears as |TI ie| (``\Red{T}arget-outside,
IFC-inside''); in the IFC language, a target language expression appears as |IT
te| ( ``{\color{blue}{I}}FC-outside, target-inside'').
    \item We reinterpret |CE| to be evaluation contexts on task lists, providing definitions for |curlyRedE| and |curlyBlueE|.  These rules only operate on the first task in the task list, which by convention is the only task executing.
    \item We reinterpret |->|, an operation on a single task, in terms of |.->|, operation on task lists.  The correspondence is simple: a task executes a step and then is rescheduled in the task list according to schedule policy |alpha|.
    Figure~\ref{fig:scheduler} defines two concrete schedulers.
    \item Finally, we define some rules for scheduling, handling sandboxing tasks (which interact with the state of the target language),
    and intermediating between the borders of the two languages.
\end{itemize}

>>>>

In Fig.~\ref{fig:ml}, we give a simple, untyped lambda calculus with
mutable references and fixpoint, prepared for combination with an
information flow control language.  The presentation is mostly standard, and utilizes Felleisen-Hieb reduction
semantics~\cite{Felleisen:1992:RRS:136293.136297} to define the
operational semantics of the system.  One peculiarity is that our language
defines an evaluation context |tE|, but, the evaluation rules have been
expressed in terms of a different evaluation context |curlyRedE|;
Here, we follow the approach of Matthews and
Findler~\cite{Matthews:2007:OSM:1190216.1190220} in order to simplify combining
semantics of multiple languages.
To derive the usual operational semantics for this language, the evaluation
context merely needs to be defined as |conf tS te @= tS, bracket tE te|.
However, when we combine this language with an IFC language, we
reinterpret the meaning of this evaluation context.

@@

In Fig.~\ref{fig:ifc}, we give the syntax and \emph{single-task}
evaluation rules for a minimal information-flow control language.
Ordinarily, information flow control languages are defined by directly
stating a base language plus information flow control operators.  In
contrast, our language is purposely minimal: it does not have sequencing
operations, control flow, or other constructs.  However, it contains
support for the following core information flow control features:

@@

\subsection{The Embedding}
Fig.~\ref{fig:embedding}
provides all of the rules responsible for actually carrying out the embedding of the IFC language with the target language.
The most important feature of this embedding is that every task maintains its own
copy of the target language global state and evaluation context, thus
enforcing isolation between various tasks.  In more detail:

@@

\begin{itemize}
    \item We extend the values, expressions and evaluation contexts of
      both languages to allow for embeddings terms in one language to
      be embedded in the other, as
      in~\cite{Matthews:2007:OSM:1190216.1190220}.
    %
In the target language, an IFC expression appears as |TI ie| (``\Red{T}arget-outside,
IFC-inside''); in the IFC language, a target language expression appears as |IT
te| ( ``{\color{blue}{I}}FC-outside, target-inside'').
    \item We reinterpret |CE| to be evaluation contexts on task lists, providing definitions for |curlyRedE| and |curlyBlueE|.  These rules only operate on the first task in the task list, which by convention is the only task executing.
    \item We reinterpret |->|, an operation on a single task, in terms of |.->|, operation on task lists.  The correspondence is simple: a task executes a step and then is rescheduled in the task list according to schedule policy |alpha|.
    Fig.~\ref{fig:scheduler} defines two concrete schedulers.
    \item Finally, we define some rules for scheduling, handling sandboxing tasks (which interact with the state of the target language),
    and intermediating between the borders of the two languages.
\end{itemize}
